---
title: "Extract_Features_5_Cells"
author: "Sizun Jiang"
date: "1/28/2019"
output: html_document
---

This markdown will do the following:
1. Extract feature counts with a sliding window across 5 cells
2. Sample across 5 cells to perform dimension reduction assays
3. Plot pretty pictures
```{r, include=FALSE, echo = FALSE}
library(bioimagetools)
library(ggplot2)
library(gplots)
# library(umap)
library(Rtsne)
library(RColorBrewer)
library(gridExtra)
library(spatstat)
source("srIBI_functions.R")
library(parallel)
library("fastcluster")
```

## Read in cells
```{r}
cells = c("Cell1", "Cell2", "Cell3", "Cell4", "Cell5")
# Load tiffs
cells_tiffs = lapply(cells, function(cell){
  tiffs = list.files(paste0(cell, "/TIFsNoBg/"), pattern = "*.tiff")
  names(tiffs) = unlist(strsplit(tiffs, split = ".tiff"))
  cell_combined = lapply(tiffs, function(tiff){
  cell_tiff = readTIF(paste0(cell, "/TIFsNoBg/", tiff), as.is = T)
})
  return(cell_combined)
})
names(cells_tiffs) = cells
```
Lets look at the images to make sure legit, summing all 20 stacks
```{r, fig.width=5, fig.height=5, dpi=300}
lapply(cells_tiffs, function(cells){
  lapply(cells, function(tiff){
  img(tiff[,,1:20], up = 1)
})
})

```

## Feature Extraction
Now lets do a pixel extraction of X by Y by Z size
Note this part is multicore
```{r}
x = 10
y = 10
z = 5
step_xy = 5
step_z = 3
discard = 100
cells_features = mclapply(cells_tiffs, function(cell_tiff){
  extract_features(cell_tiff, x = x, y = y, z = z, step_xy = step_xy, step_z = step_z, discard = discard)
}, mc.cores = 8)
```

Sanity check. Plots all the features
```{r, fig.width=5, fig.height=5, dpi=300}
dim_xy = floor((1024-x)/step_xy)

lapply(cells_features, function(ext_features){
  test_df = ext_features$`19F-Nucleolin`
  img(t(matrix(test_df$`19F-Nucleolin`[test_df$z_start == 1], 
               sqrt(length(test_df$`19F-Nucleolin`[test_df$z_start == 1])),
               sqrt(length(test_df$`19F-Nucleolin`[test_df$z_start == 1])))))
})
```
Concatenate DFs and get log2
```{r}
cells_log2_all_features = lapply(cells_features, function(ext_features){
  all_features = do.call(cbind, lapply(ext_features, function(feat){
  return(feat[,2])
  }))
  log2_all_features_df = cbind(ext_features[[1]][,-2], log2(all_features+0.0001))
})
```

# Sample and cluster
```{r}
set.seed(123456)
# Number to sample per cell
sample_num = 20000
subset_cells = do.call(rbind, lapply(names(cells_log2_all_features), function(cell){
   sample_cells(cells_log2_all_features[[cell]], sample_num = sample_num, cellID = cell)
}))


# Use all channels but 12C and SED
data.tsne = Rtsne(subset_cells[,c(8, 10:14)], check_duplicates = F, num_threads = 8)
df.tsne.input = (subset_cells[,c(8, 10:14)])
df.tsne = as.data.frame(data.tsne$Y)
# Plot
plot(data.tsne$Y,asp=1) # Plot the result
```

# Clustering of the pixels/boxes (131220 voxels per cell)
# Note that you'll need a lot of ram for this
```{r}
mycolors = c(brewer.pal(name = "Set1", 9), brewer.pal(name="Set3", 12), brewer.pal(name="Set2", 8))
df.tsne.clust = df.tsne
cluster_k_results = lapply(3:30, function(k){
  return(kmeans(scale(df.tsne), k))
})
names(cluster_k_results) = 3:30
cluster_h_results = hclust(dist(scale(df.tsne)))
```



```{r, fig.width=15, fig.height=8, dpi=300}
for(k in 3:24){
  df.tsne.clust$cl_kmeans = factor(cluster_k_results[[as.character(k)]][["cluster"]])
  df.tsne.clust$cl_hierarchical = factor(cutree(cluster_h_results, k=k))
  plot_k1=plot_cluster_multicolor(df.tsne.clust, "cl_kmeans", mycolors)
  plot_h1=plot_cluster_multicolor(df.tsne.clust, "cl_hierarchical", mycolors)
  grid.arrange(plot_k1, plot_h1, ncol=2)
}
```
Lets color the tSNE map by expression
```{r,  fig.width=8, fig.height=12, dpi=300}
# Channel info
df_channels = do.call(cbind, lapply(colnames(df.tsne.input[1:6]), function(x){
  ch = scale(df.tsne.input[[x]], center = T)
  return(ch)
}))
colnames(df_channels) = c("H3K27ac", "Cisplatin", "SC35", "Nucleolin", "DNA", "H3K9me3")
df.tsne.clust_range = cbind(df.tsne.clust, df_channels)
```


Extract cluster info per cluster. Median
```{r, fig.width=8, fig.height=5, dpi=300}

for(k in 29:29){
  df.tsne.input$cl_hierarachical = factor(cutree(cluster_h_results, k=k))
  hierarachical_ave = aggregate(df.tsne.input[1:6], list(df.tsne.input$cl_hierarachical), FUN = median)
  require(gplots)
  col = colorRampPalette(c("blue","white","red"))(256)
  hm_h = heatmap.2(t(as.matrix(hierarachical_ave[,2:7])), scale = "row", col = col,
            sepwidth=c(0.01,0.01), sepcolor = "black", colsep=1:ncol(t(as.matrix(hierarachical_ave[,2:7]))),rowsep=1:nrow(t(as.matrix(hierarachical_ave[,2:7]))),
            density.info="none", trace="none", margins=c(3,12), main = paste0("Hierarchical Clustering, k =", k))
}
```

Use mean here
```{r, fig.width=8, fig.height=5, dpi=300}

for(k in 29:29){
  df.tsne.input$cl_hierarachical = factor(cutree(cluster_h_results, k=k))
  hierarachical_ave = aggregate(df.tsne.input[1:6], list(df.tsne.input$cl_hierarachical), FUN = mean)
  require(gplots)
  col = colorRampPalette(c("blue","white","red"))(256)
  hm_h = heatmap.2(t(as.matrix(hierarachical_ave[,2:7])), scale = "row", col = col,
            sepwidth=c(0.01,0.01), sepcolor = "black", colsep=1:ncol(t(as.matrix(hierarachical_ave[,2:7]))),rowsep=1:nrow(t(as.matrix(hierarachical_ave[,2:7]))),
            density.info="none", trace="none", margins=c(3,12), main = paste0("Hierarchical Clustering, k =", k))
}
```

| Mean, k = 29             |                                   |                 |
| ------------------------ | --------------------------------- | --------------- |
| Annotation               | Clusters                          | New cluster Num |
| Heterochromatin 1        | 1, 5                              | 1               |
| Heterochromatin 2        | 13, 27                            | 2               |
| Euchromatin 1            | 11                                | 3               |
| Euchromatin 2            | 16, 19                            | 4               |
| Euchromatin + cisplatin  | 3, 18                             | 5               |
| SC35 + Ac + Cisplatin    | 4, 7, 10                          | 6               |
| SC35, me3 + Ac           | 25                                | 7               |
| SC35                     | 6, 9                              | 8               |
| Nucleolin low cisplatin  | 2, 12, 14, 15, 17, 23, 24, 26, 28 | 9               |
| Nucleolin high cisplatin | 22                                | 10              |
| Toss                     | 20, 21, 29                        | 11              |

1. Heat maps after combining
2. Tsne after combining

```{r}
k = 29
df.tsne.input$cl_hierarachical = factor(cutree(cluster_h_results, k=k))
new_cl_hierarachical = factor(cutree(cluster_h_results, k=k))
w1 = which(new_cl_hierarachical == 1| new_cl_hierarachical == 5)
w2 = which(new_cl_hierarachical == 13| new_cl_hierarachical == 27)
w3 = which(new_cl_hierarachical == 11)
w4 = which(new_cl_hierarachical == 16| new_cl_hierarachical == 19)
w5 = which(new_cl_hierarachical == 3| new_cl_hierarachical == 18)
w6 = which(new_cl_hierarachical == 4| new_cl_hierarachical == 7| new_cl_hierarachical == 10)
w7 = which(new_cl_hierarachical == 25)
w8 = which(new_cl_hierarachical == 6| new_cl_hierarachical == 9)
w9 = which(new_cl_hierarachical == 2| new_cl_hierarachical == 12| new_cl_hierarachical == 14| new_cl_hierarachical == 15| new_cl_hierarachical == 17| new_cl_hierarachical == 23| new_cl_hierarachical == 24| new_cl_hierarachical == 26| new_cl_hierarachical == 28)
w10 = which(new_cl_hierarachical == 22)
w11 = which(new_cl_hierarachical == 20| new_cl_hierarachical == 21| new_cl_hierarachical == 29)
new_cl_hierarachical[w1] = 1
new_cl_hierarachical[w2] = 2
new_cl_hierarachical[w3] = 3
new_cl_hierarachical[w4] = 4
new_cl_hierarachical[w5] = 5
new_cl_hierarachical[w6] = 6
new_cl_hierarachical[w7] = 7
new_cl_hierarachical[w8] = 8
new_cl_hierarachical[w9] = 9
new_cl_hierarachical[w10] = 10
new_cl_hierarachical[w11] = 11
df.tsne.input$new_cl_hierarachical = factor(new_cl_hierarachical)
table(df.tsne.input$new_cl_hierarachical)

hierarachical_median = aggregate(df.tsne.input[1:6], list(df.tsne.input$new_cl_hierarachical), FUN = mean)
col = colorRampPalette(c("blue","white","red"))(256)
hm_h = heatmap.2(t(as.matrix(hierarachical_median[,2:7])), scale = "row", col = col,
          sepwidth=c(0.01,0.01), sepcolor = "black", colsep=1:ncol(t(as.matrix(hierarachical_median[,2:7]))),rowsep=1:nrow(t(as.matrix(hierarachical_median[,2:7]))),
          density.info="none", trace="none", margins=c(3,12), main = paste0("Hierarchical Clustering"))
```

Plot combined clusters
```{r, fig.width=12, fig.height=15, dpi=300}
df.tsne.clust$cl_hierarachical = df.tsne.input$cl_hierarachical
df.tsne.clust$new_cl_hierarachical = df.tsne.input$new_cl_hierarachical
plot_h_old=plot_cluster_multicolor(df.tsne.clust, "cl_hierarachical", mycolors)
plot_h_combined=plot_cluster_multicolor(df.tsne.clust, "new_cl_hierarachical", mycolors)
grid.arrange(plot_gradient(df.tsne.clust_range , "H3K27ac"), 
           plot_gradient(df.tsne.clust_range, "Cisplatin"),
           plot_gradient(df.tsne.clust_range, "SC35"),
           plot_gradient(df.tsne.clust_range, "Nucleolin"),
           plot_gradient(df.tsne.clust_range, "DNA"),
           plot_gradient(df.tsne.clust_range, "H3K9me3"),
           plot_h_old,
           plot_h_combined,
           ncol=3, nrow=3)
```


```{r, fig.width=8, fig.height=9, dpi=300}
mycolors2 = c(brewer.pal(name = "Set3", 11 ), brewer.pal(name = "Set1", 9)[7])
plot_cluster_multicolor(df.tsne.clust, "new_cl_hierarachical", mycolors2)
```

Permutation test implemented for more stringent testing of neighborhoods:

Pairwise interactions between and within cell phenotypes are compared to a random distribution using two individual one-tailed permutation tests (equation (1)). These provide two P values which correspond to either interaction or avoidance. These P values represent the likelihood of a neighborhood interaction being enriched or depleted in comparison to a randomized version of the same tissue. The comparison to a matched randomized tissue for every individual image controls for both the distinct connectivity and the specific cell types in that tissue. Equation (1) describes our approach using a permutation test with Monte Carlo sampling. We run this test twice to calculate the P value for each tail.

Nearest Neighbor Permutation Test
```{r}
subset_cells$h_cluster = df.tsne.input$new_cl_hierarachical
# Permutation test

perm_6_10 = perm_AB(subset_cells, namA=6, namB=10, radius=5, n=1000, mc.cores = 38)
# Plot for permutation test between for cluster 6's assocaition with cluster 10
lapply(cells, function(cell){
  perm_6_10$plots[[cell]]
})

# Pvalue that interaction is far lower than expected by random
perm_6_10$adj.pval
# Log2 Fold Change over mean of perm test
perm_6_10$log2FC
```
Perform a pairwise test for all 11 clusters against each other. Try diff rad
Done:5, 25
Doing: for c(1,2,5,10,25,50)
```{r}
clusters = sort(unique(subset_cells$h_cluster), decreasing = F)
names(clusters) = clusters
pairwise_perm_test_all = lapply(c(1,2,5,10,25,50), function(i){
  print(i)
  pairwise_perm_test = lapply(clusters, function(A){
  AtoB_perm_test = mclapply(clusters, function(B){
    perm_AB(subset_cells, namA=A, namB=B, radius=i, n=1000, mc.cores = 20)
    })
  })
})
# Names is radius
names(pairwise_perm_test_all) = c(1,2,5,10,25,50)


# pairwise_perm_test_5rad_named = lapply(pairwise_perm_test_5rad, function(clust){
#   names(clust) = clusters
#   return(clust)
# })

save(pairwise_perm_test_all, file = "pairwise_perm_test_all.list")
```

# See if there is a trend in different radius
c(1,2,5,10,20,25,30,50,100)
```{r, fig.width=8, fig.height=6, dpi=300}
pairwise_plots = lapply(names(pairwise_perm_test_all), function(x){
  perm_matrix = matrix(nrow = length(clusters), ncol = length(clusters))
  for(i in 1:length(clusters)){
    for(j in 1:length(clusters)){
      perm_matrix[i,j] = pairwise_perm_test_all[[x]][[i]][[j]]$log2FC
    }
  }
# No clustering
heatmap.2(perm_matrix, scale = "row", col = col, Rowv = F, Colv = F,
          sepwidth=c(0.01,0.01), sepcolor = "black", colsep=1:ncol(perm_matrix[,1:11]),rowsep=1:nrow(perm_matrix[,1:11]),
          density.info="none", trace="none", margins=c(3,12), main = paste0("Interaction Frequency Radius = ", x),
          key.xlab="Log2FC")
  
})

```

Plot Heatmap rad = 5
```{r, fig.width=8, fig.height=6, dpi=300}
perm_matrix = matrix(nrow = length(clusters), ncol = length(clusters))
for(i in 1:length(clusters)){
  for(j in 1:length(clusters)){
    perm_matrix[i,j] = pairwise_perm_test_5rad_named[[i]][[j]]$log2FC
  }
}

# No clustering
# heatmap.2(perm_matrix, Rowv = F, Colv = F, dendrogram = "none", symm = T)
heatmap.2(perm_matrix, scale = "row", col = col, Rowv = F, Colv = F,
          sepwidth=c(0.01,0.01), sepcolor = "black", colsep=1:ncol(perm_matrix[,1:11]),rowsep=1:nrow(perm_matrix[,1:11]),
          density.info="none", trace="none", margins=c(3,12), main = paste0("Interaction Frequency"),
          key.xlab="Log2FC")

# Clustered
heatmap.2(perm_matrix, scale = "row", col = col, Rowv = T, symm=T,
          sepwidth=c(0.01,0.01), sepcolor = "black", colsep=1:ncol(perm_matrix[,1:11]),rowsep=1:nrow(perm_matrix[,1:11]),
          density.info="none", trace="none", margins=c(3,12), main = paste0("Interaction Frequency"),
          key.xlab="Log2FC")

# Get matrix of pvalues, mean adjusted P
pval_matrix = matrix(nrow = length(clusters), ncol = length(clusters))
for(i in 1:length(clusters)){
  for(j in 1:length(clusters)){
    pval_matrix[i,j] = mean(pairwise_perm_test_5rad_named[[i]][[j]]$adj.pval)
  }
}

pval_matrix_binary = pval_matrix
pval_matrix_binary= (pval_matrix_binary < 0.05)*1
heatmap.2(pval_matrix_binary, scale = NULL, Rowv = F, symm=T,
          sepwidth=c(0.01,0.01), sepcolor = "black", colsep=1:ncol(pval_matrix[,1:11]),rowsep=1:nrow(pval_matrix[,1:11]),
          density.info="none", trace="none", margins=c(3,12), main = paste0("Pval of Interaction Frequency"),
          key.xlab="Significant (1) or Not (0)")

```

Plot Heatmap rad = 25
```{r, fig.width=8, fig.height=6, dpi=300}
perm_matrix = matrix(nrow = length(clusters), ncol = length(clusters))
for(i in 1:length(clusters)){
  for(j in 1:length(clusters)){
    perm_matrix[i,j] = pairwise_perm_test_25rad[[i]][[j]]$log2FC
  }
}

# No clustering
# heatmap.2(perm_matrix, Rowv = F, Colv = F, dendrogram = "none", symm = T)
heatmap.2(perm_matrix, scale = "row", col = col, Rowv = F, Colv = F,
          sepwidth=c(0.01,0.01), sepcolor = "black", colsep=1:ncol(perm_matrix[,1:11]),rowsep=1:nrow(perm_matrix[,1:11]),
          density.info="none", trace="none", margins=c(3,12), main = paste0("Interaction Frequency"),
          key.xlab="Log2FC")

# Clustered
heatmap.2(perm_matrix, scale = "row", col = col, Rowv = T, symm=T,
          sepwidth=c(0.01,0.01), sepcolor = "black", colsep=1:ncol(perm_matrix[,1:11]),rowsep=1:nrow(perm_matrix[,1:11]),
          density.info="none", trace="none", margins=c(3,12), main = paste0("Interaction Frequency"),
          key.xlab="Log2FC")

# Get matrix of pvalues, mean adjusted P
pval_matrix = matrix(nrow = length(clusters), ncol = length(clusters))
for(i in 1:length(clusters)){
  for(j in 1:length(clusters)){
    pval_matrix[i,j] = mean(pairwise_perm_test_25rad[[i]][[j]]$adj.pval)
  }
}

pval_matrix_binary = pval_matrix
pval_matrix_binary= (pval_matrix_binary < 0.05)*1
heatmap.2(pval_matrix_binary, scale = NULL, Rowv = F, symm=T,
          sepwidth=c(0.01,0.01), sepcolor = "black", colsep=1:ncol(pval_matrix[,1:11]),rowsep=1:nrow(pval_matrix[,1:11]),
          density.info="none", trace="none", margins=c(3,12), main = paste0("Pval of Interaction Frequency"),
          key.xlab="Significant (1) or Not (0)")

```



```{r}
channel_plots = lapply(colnames(subset_cells)[8:15], function(i){
  plot_marker_dist(cluster_list = cluster_list, channel_name = i)
})
names(channel_plots) = colnames(subset_cells)[8:15]

channel_plots[[1]]
channel_plots[[2]]
channel_plots[[3]]
channel_plots[[4]]
channel_plots[[5]]
channel_plots[[6]]
channel_plots[[7]]
# plot_marker_dist(cluster_list = cluster_list, channel_name = '194Pt-Cisplatin', whisker_cutoff = 1.05)
```